\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-10-08 icra-2021.tex
paper fixed with latests

The algorithm inputs the mission
\begin{equation}
  \mathcal{M}:=\left\{\varphi_k,\,(\underline{c}_k,\overline{c}_k),\bigcup_{i=1,\dots,\rho}(\underline{\text{QoS}}_{k,i},\overline{\text{QoS}}_{k,i}):\,\forall k\in\mathbb{T}\right\},
\end{equation}
where $\mathbb{T}:=\,[0,t_f)\subseteq\mathbb{R}_{\geq 0}$ is the mission time, $t_f$ the final instant, $\varphi_k\in\mathbb{R}$ is a generic continuous twice differentiable TEE along its upper and lower bounds $(\underline{c}_k,\overline{c}_k)\in\mathbb{R}^2$ (the motion plan), and a set of upper and lower bounds for the $\sigma$ tasks $(\underline{\text{QoS}}_{k,i},\overline{\text{QoS}}_{k,i})\in\mathbb{Z}_{\geq 0}^2$.

@@

It outputs the correction to the position and the instantaneous energy consumption, while it adapts the controls $\mathbf{u}_k:=(c_k,\text{QoS}_{k,1},\dots,\text{QoS}_{k,\rho})$, the TEE and QoS parameters.

@@

\subsection{Position evolution}
\label{sec:position}

@@

Let us define $\mathcal{P}_k$, the set that delimits all the possible deviations form the $k$-th TEE
\begin{equation}\label{eq:area}
  \mathcal{P}_k:=\{\mathbf{p}:\underline{c}_k\leq\varphi_k(\mathbf{p},c_{k,1},\dots,c_{k,\rho})\leq\overline{c}_k\},
\end{equation}
where $\underline{c}_k,\overline{c}_k$ are the bounds where the TEE $\varphi_k:\mathbb{R}^2\times\mathbb{R}^{\sigma}\rightarrow\mathbb{R}$ is free to evolve using controls $\mathbf{c}_k:=c_{k,1},...,c_{k,\sigma}$.

@@

The energy evolution analysis necessitates the following realistic assumption.
\begin{assm}\label{assm:periodic} The position evolution is periodic, in the sense that it presents repetitive patterns.
\end{assm}
Formally, given two time instants $k_1,k_2\in\mathbb{R}_{\geq 0}$ such that $k_1>k_2$ and a value $n\in\mathbb{R}_{\geq 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}$ such that
\begin{equation}
  \|\mathbf{p}(t)-\mathbf{p}(t+n)\|= e\,\,\,\forall t\in\{k_1,...,k_2\}.
\end{equation}
We show in Section~\ref{sec:experimental} the assumption being eased in practice to a set $\mathbb{E}\subset\mathbb{R}$.

Let us consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$
\begin{equation}\label{eq:fourier}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients.

The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)\\%+w(t)\\
  y(t)&=C\mathbf{q}(t)%+v(t)
\end{cases},\end{equation}
where $y(t)\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. We prove formally in the Theorem~\ref{thm:state-vs-energy} the instantaneous energy being obtained as a linear combination of the state. The control $\mathbf{u}$ along with the input matrix $B$ are defined later in Subsection~\ref{sec:control-action}, the state $\mathbf{q}$ mimics the original Fourier series coefficients, and

\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}(t)\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. In matrix $A$, the first value is one, $A_n$ is later on the diagonal, and zero in the remaining.

We further model the system of Equation~(\ref{eq:state-perf}) plus disturbances with the following discrete time model
\begin{equation}\label{eq:state-unct}
  \begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
  \end{cases},
\end{equation}
where $\mathbf{w}_k\in\mathbb{R}^j,v_k\in\mathbb{R}$ account for unknown state and output disturbances. In the case of the system behaving ideally (i.e., with zero disturbances), we expect a set of states evolving accordingly to their output. Such observation is summarized in the following Lemma.

@@

The energy cost of the computations is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds an energy model. Specifically, the tool builds a multivariate linear interpolation from the mission specification. It hence requires the user to implement a ROS system with one or more computationally expensive ROS nodes which change the computational load by node-specific ROS parameters (QoS).

Suppose the system is composed of $\sigma$ computationally expensive ROS nodes. Let us define the computational control action
\begin{equation}\label{eq:qos-def}
  \mathcal{C}_k:=\left\{u:u\in\text{QoS}_n(k)\,\,\,\forall n\in\{0,\dots,\sigma-1\}\right\},
\end{equation}
where $\text{QoS}_n(k):\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $n$-th QoS value at time instant $k$, and $\mathcal{C}_k\in\mathbb{Z}_{\geq 0}^\sigma$ the set of $\sigma$ QoS values the system is composed of. Let us further define $g:\mathbb{Z}_{\geq 0}\times\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous energy value obtained interrogating \powprof{}. The instantaneous computational energy component can be defined
\begin{equation}\label{eq:energy-comp}\begin{split}
  y_k^c:&=g\left(\text{QoS}_0\left(k\right),\dots,\text{QoS}_{\sigma-1}\left(k\right)\right)=g\left(\mathcal{C}_{k}\right).\\
\end{split}\end{equation}

The QoS parameters $\mathcal{C}_k$ can be subject to different constraints at different states. Physically, this means that the UAV can perform the ROS nodes within different QoS ranges in time while flying
\begin{equation}\label{eq:qos-lims}\begin{split}
  \underline{\text{QoS}}_n(k)\leq \,&\text{QoS}_n(k)\leq \overline{\text{QoS}}_n(k),\\
  &\forall n\in\{0,\cdots,\sigma-1\},
\end{split}\end{equation}
where the values $\underline{\text{QoS}}_n(k),\overline{\text{QoS}}_n(k)$ are retrieved from the mission specification.

\begin{center}\vspace{.2cm}---\emph{I'll move this subsec under eq 9, it doesn't make sense now and my brain stopped working for this evening}---\vspace{.6cm}\end{center}

\subsection{Control action}
\label{sec:control-action}

We build the control action $\mathbf{u}$ from Equation~(\ref{eq:state-perf})~and~(\ref{eq:state-unct}) in two steps. In Equation~(\ref{eq:qos-def}) we defined the control due to the computations, here we do the same for the trajectory.

Let us re-define the TEE introduced first in Subsection~\ref{sec:model} as a function $\varphi_k:\mathbb{R}^3\times\mathbb{R}^\rho\rightarrow\mathbb{R}$ of the position and $\rho$ TEE parameters $\mathcal{M}\in\mathbb{R}^\rho$, e.g., the constants of a linear function, the radius of a circle, and semi-major and minor axis of an ellipse. Given $\mathbf{u}_{k}:=\{u_{k,0},\dots,u_{k,\rho-1}\}$, the set of these parameters at time instant $k$ can be expressed
\begin{equation}\label{eq:tee-def}
  \mathcal{M}_k:=\{u_k\in\mathbf{u}_k : \varphi_k(\mathbf{p},\mathbf{u}_k)\in\mathcal{P}_k\},
\end{equation}
where $\mathbf{p}\in\mathcal{P}_k$ is any optimal point which let the trajectory explicit function $\varphi_k$ converge under any optimal TEE parameter $u_k^0$. Physically, $\mathcal{M}_k$ contains all the controls generating the points over the area $\mathcal{P}$ defined in Equation~(\ref{eq:area}).

The explicit trajectory equation $\varphi_k$ can be different at different states $k$, meaning the vector field and guidance action, from Equation~(\ref{eq:pd})~and~(\ref{eq:uav-cont}) respectively, will account for the sudden change of trajectory during the mission.

It is worth considering that the number of parameters at state $k$ is a parameter of the state. This is for the sake of generality, as the mission specification might contain different explicit equations for different states. For instance, the fixed-wing craft might follow an ellipse function throughout the mission and heading a linear function while landing.

The TEE parameters and QoS, $\mathcal{M}$ and $\mathcal{C}$ defined in Equation~(\ref{eq:tee-def})~and~(\ref{eq:qos-def}), are incorporated in the system in Equation~(\ref{eq:state-unct}) using the input matrix
\begin{equation}
  \mathbf{u}_k=\left[\begin{array}{c}
    g\left(\mathcal{C}_k\right)-g\left(\mathcal{C}_{k-1}\right) \\\hline \mathcal{M}_k
  \end{array}\right],\,\,\,
  B=\left[\begin{array}{ccccccc}
    1 & \vline & 0 & \cdots & 0\\\hline
    0 & \vline & 0 & \cdots & 0\\
    \vdots & \vline & \vdots & \ddots & \vdots\\
    0 & \vline & 0 & \cdots & 0
  \end{array}\right],
\end{equation}
where $\mathbf{u}_k\in\mathbb{R}^l$ is the control given $l:=1+\rho$, and $B\in\mathbb{R}^{j\times l}$ is the input matrix from Equation~(\ref{eq:state-perf})~and~(\ref{eq:state-unct}). Moreover, the first column in the first row of the input matrix is $1$, while all the other items are $0$. This adds the computational model component to the energy evolution in the system. The energy due to the change of explicit trajectory equation parameters is not directly added to the system, which however will update the reading from the sensors $y^s$ defined in Subsection~\ref{sec:state-est} and adjust the energy evolution accordingly.

@@

The main goal of the algorithm is to find an optimal control action $\mathbf{u}^0$ for the current state $\hat{\mathbf{q}}$--estimated from sensors' measurements--from the optimal control law $\mathbf{u}^0=:\kappa(\hat{\mathbf{q}})$. This is achieved by solving online a finite horizon optimal control problem by the hand of a modification of a model predictive control (MPC) algorithm~\cite{rawlings2017model}.

>>>>

The algorithm inputs the mission
\begin{equation}\label{eq:mission}
  \mathcal{M}:=\left\{\varphi_k,\,(\underline{c}_k,\overline{c}_k),\bigcup_{n=0,\dots,\rho-1}(\underline{\text{QoS}}_{k,n},\overline{\text{QoS}}_{k,n}):\,\forall k\in\mathbb{T}\right\},
\end{equation}
where $\mathbb{T}:=\,[0,t_f)\subseteq\mathbb{R}_{\geq 0}$ is the mission time, $t_f$ the final instant, $\varphi_k\in\mathbb{R}$ is a generic continuous twice differentiable TEE along its upper and lower bounds $(\underline{c}_k,\overline{c}_k)\in\mathbb{R}^2$ (the motion plan), and a set of upper and lower bounds for the $\sigma$ tasks $(\underline{\text{QoS}}_{k,i},\overline{\text{QoS}}_{k,i})\in\mathbb{Z}_{\geq 0}^2$.

@@

It outputs the correction to the position and the instantaneous energy consumption, while it adapts the control $\mathbf{u}_k$ in function of the TEE and QoS parameters.

@@

\subsection{Trajectory}
\label{sec:position}

@@

Let us define $\mathcal{P}_k$, the set that delimits all the possible deviations form the $k$-th TEE
\begin{equation}\label{eq:area}
  \mathcal{P}_k:=\{\mathbf{p}:\underline{c}_k\leq\varphi_k(\mathbf{p},c_{k,1},\dots,c_{k,\sigma})\leq\overline{c}_k\},
\end{equation}
where $\underline{c}_k,\overline{c}_k$ are the bounds from $\mathcal{M}$ that disclose the are over which the TEE $\varphi_k:\mathbb{R}^2\times\mathbb{R}^{\sigma}\rightarrow\mathbb{R}$ is free to evolve using controls $\mathbf{c}_k:=c_{k,1},...,c_{k,\sigma}$.

@@

The energy evolution analysis necessitates the following realistic assumption.
\begin{assm}\label{assm:periodic} The trajectory is periodic, in the sense that it presents repetitive patterns.
\end{assm}
Formally, given two time instants $k_1,k_2\in\mathbb{T}$ s.t. $k_1>k_2$ and a value $n\in\mathbb{R}_{\geq 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}$
\begin{equation}
  \|\mathbf{p}_{k}-\mathbf{p}_{k+n}\|= e\,\,\,\forall k\in[k_1,k_2].
\end{equation}
We show in Section~\ref{sec:experimental} the assumption being eased in practice to a set $\mathbb{E}\subset\mathbb{R}$ or omitted under specific conditions.

Let us consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of energy estimation of such a periodic trajectory
\begin{equation}\label{eq:fourier}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients.

Suppose no disturbances in form of $\mathbf{w}_k\in\mathbb{R}^j,v_k\in\mathbb{R}$ accounting otherwise for the unknown state and output. The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
\end{cases},\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. We prove formally in the Theorem~\ref{thm:state-vs-energy} the instantaneous energy being obtained as a linear combination of the state. The state $\mathbf{q}$ mimics the original Fourier series coefficients, and
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. In matrix $A$, the first value is one, $A_n$ is later on the diagonal, and zero in the remaining.

The control $\mathbf{u}$ along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\text{QoS}_k)-g(\text{QoS}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T\\
  B&=\left[\begin{array}{cccc}
    1&         &       &  \\
     & \omega_{k,1}&       &  \\
     &         & \ddots&  \\
     &         &       & \omega_{k,\sigma}
  \end{array}\right]
\end{split}\end{equation}
where $\text{QoS}_k$ is defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^l$ is the control given $l:=1+\rho$, $\mathbf{u}_{-1}=\begin{bmatrix}0 & \cdots & 0\end{bmatrix}$ and $B\in\mathbb{R}^{j\times l}$. Moreover, the first item on the diagonal is one, while the others are gain factors $\omega_{k}\in\mathbb{R}$, quantifying the contribution of a given TEE parameter to the instantaneous energy. Equation~(\ref{eq:state-control}) accounts for the energy due to the computations. The energy due to the change of explicit trajectory equation parameters is merely a linear combination of the gain factor and the value of the TEE. Nevertheless, the change updates the trajectory which will hence affect the reading from the sensors and adjust the energy evolution accordingly.

In the case of the system behaving ideally (i.e., with zero disturbances), we expect a set of states evolving accordingly to their outputs. Such observation is summarized in the following Lemma.

@@

The energy cost of the computations is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds an energy model. Specifically, the tool builds a linear interpolation, one per each task. It hence requires the user to implement the mission as a ROS system with one or more ROS nodes which change the computational load by node-specific ROS parameters (QoS).

Suppose the system is composed of $\sigma$ ROS nodes. Let us define a computational control action
\begin{equation}\label{eq:qos-def}
  \mathcal{C}_k:=\left\{u:u\in\text{QoS}_{k,n}\,\,\,\forall n\in\{0,\dots,\sigma-1\},\,\forall k\in\mathbb{T}\right\},
\end{equation}
where $\text{QoS}_n(k):\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $n$-th QoS value at time instant $k$, and $\mathcal{C}_k\in\mathbb{Z}_{\geq 0}^\sigma$ the set of $\sigma$ QoS values. Let us further define $g:\mathbb{Z}_{\geq 0}\times\mathbb{Z}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous energy value obtained interrogating \powprof{}. The instantaneous computational energy component can be defined
\begin{equation}\label{eq:energy-comp}\begin{split}
  y_k^c:&=g\left(\text{QoS}_{k,0},\dots,\text{QoS}_{k,\sigma-1}\right)=g\left(\mathcal{C}_{k}\right).\\
\end{split}\end{equation}

The QoS parameters $\mathcal{C}_k$ can be subject to different constraints at different states. Physically, this means that the UAV can perform the ROS nodes within different QoS ranges in time while flying
\begin{equation}\label{eq:qos-lims}\begin{split}
  \underline{\text{QoS}}_{k,n}\leq \,&\text{QoS}_{k,n}\leq \overline{\text{QoS}}_{k,n},\\
  &\forall n\in\{0,\cdots,\sigma-1\},\,\forall k\in\mathbb{T},
\end{split}\end{equation}
where the values $\underline{\text{QoS}}_{k,n},\overline{\text{QoS}}_{k,n}$ are retrieved from the mission $\mathcal{M}$.

@@

%The main goal of the algorithm is to find an optimal control action $\mathbf{u}^0$ for the current state $\hat{\mathbf{q}}$--estimated from sensors' measurements--from the optimal control law $\mathbf{u}^0=:\kappa(\hat{\mathbf{q}})$. This is achieved by solving online a finite horizon optimal control problem by the hand of a modification of a model predictive control (MPC) algorithm~\cite{rawlings2017model}.
